---
title: "Tracking quality results"
output: html_notebook
---

```{r setup}

library(tidyverse)
library(nloptr)
library(ggpubr)
library(zoo)

lindf <- readRDS("lindf.rds")

sbj_cols <- ggsci::pal_d3("category10")(10)[c(1,2,4:10)]

update_geom_defaults("point", list(shape = 16))

```


# Fig 1: example trial

same color coding (pre-launch, preview, occlusion) in other figures?

```{r}
# position
(p <- lindf %>%
  filter(participant == 7) %>%
  filter(tn_linear == 64) %>% 
  filter(trial_ts > -.5, trial_ts < 1.129982686 + .05) %>% 
   mutate(block = factor(block)) %>%
  arrange(trial_ts) %>%
   gather(var, value, gX, target_x) %>%
   mutate(var = recode(var, target_x = "Object", gX = "Gaze")) %>%
  ggplot() + 
  # annotate("rect", xmin = 0.365990878, xmax = 1.129982686, ymin = -Inf, ymax = Inf,  alpha = .2) +
#  geom_rect(aes(xmin = 0.365990878, xmax = 1.129982686, ymin = -Inf, ymax = Inf), alpha = .005) + 
  geom_line(aes(trial_ts, value + 26.5, linetype = var), size=.6) +
  theme_classic(base_size = 14) +
  labs(x = "", y = "Horizontal position (deg)", linetype = "") +
  scale_x_continuous(breaks = c(-.5, 0, .5, 1.0)) +
  xlim(-.5,1.3) +
  theme(legend.position = c(.85,.3),
        plot.margin = unit(c(0, 0, 0, 0), 'lines'),
        axis.text.x = element_blank()))


(p2 <- p  + 
  geom_bracket(xmin = -.5, xmax = -0.005, y.position = 10,
    label = "pre-launch 1s", tip.length = c(0, .05), color = "royalblue4", label.size=4.2) + 
  geom_bracket(xmin = 0, xmax = 0.365990878, y.position = 20,
    label = "preview 0.2-1s", tip.length = c(0.07, 0.02), color = "#5c5c5c", label.size=4.2)  + 
  geom_bracket(xmin = 0.365990878, xmax = 1.12, y.position = 40.8,
    label = "occlusion 0-1.9s", tip.length = c(0.07, 0.015), color = "orangered3", label.size=4.2, hjust=.7)  + 
  geom_bracket(xmin = 1.129982686, xmax = 1.129982686+.05, y.position = 41.1,
    label = "target 0.05s", tip.length = .01, color = "#5c5c5c", label.size=4.18,hjust=.66))

(p3 <- p2 + annotate("text", x = -.19, y = 6, label = "anticipatory pursuit", color = "royalblue4") +
  annotate("curve", x = -.20, y = 5, xend = -.18, yend = 1.3,
    arrow = arrow(length = unit(0.08, "inch")), #size = 0.2,
    color = "royalblue4", curvature = 0) + 
  annotate("text", x = .43, y = 4.2, label = "catch-up\nsaccade", color = "#5c5c5c") +
  annotate("curve", x = .34, y = 4, xend = .22, yend = 4.8,
    arrow = arrow(length = unit(0.08, "inch")), #size = 0.2,
    color = "#5c5c5c", curvature = 0))



# gain

(p4 <- lindf %>%
  filter(participant == 7, block != 1) %>%
  filter(trial_ts > -.5, trial_ts < 1.129982686 + .05, target_class %in% c("visible", "occluded")) %>% 
   mutate(block = factor(block)) %>%
  arrange(trial_ts) %>%
  mutate(gain = gaze_speed_2/target_speed) %>%
  mutate(target_class = str_to_title(target_class)) %>%
  ggplot() + 
  geom_line(aes(trial_ts, gain, color = target_class %>% fct_rev, group = tn_linear), size=.3, alpha = .2) +
  geom_hline(yintercept=1, linetype = "dashed", alpha = .6) +
  theme_classic(base_size = 14) +
  labs(x = "Time from launch (s)", y = "Gain", color = "") +
  scale_x_continuous(breaks = c(-.5, 0, .5, 1.0), limits = c(-.5,1.3)) +
  scale_color_manual(values = c("orangered3", "#5c5c5c") %>% rev) +
  ylim(0,6) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  theme(legend.position = c(.2, .7),
        plot.margin = unit(c(0, 0, 0, 0), 'lines')))


(trial_both <- ggarrange(p3, p4, ncol = 1, heights = c(1, .6), 
               align = "hv", common.legend = F)) 

```

# Fig 2

- clarify labelling: separate axis labels (time from launch (s) similar to Fig1, and time from occlusion onset). Facet labels: Target visible, target occluded? (also for fig1 and fig3? ("visually guided" is perhaps not good; better just to say what we know i.e. target is visible or not.))
- align the axis scales
- fix smoothing of means


```{r}
# # Need to make timestamps regular for the mean lines.
new_sr = 1/120
  
resample <- function(trialdata, start_time) { # start_time = -.5 for visible, 0 for occluded
  # irregular time points at which data was sampled
  t <- trialdata$cond_ts
  # measurements 
  y <- trialdata$pos_displ
  y2 <- trialdata$speed_diff

  f <- approxfun(t,y)
  f2 <- approxfun(t,y2)
  
  # new regular data frame
  data.frame(newts = seq(from=start_time, 
                         to=max(trialdata$cond_ts), by=new_sr)) %>%
  mutate(pos_displ = f(newts),
         speed_diff = f2(newts)) 

}


ddata <- lindf %>%
  filter((target_class %in% c("visible", "occluded")) | (target_class == "other" & between(trial_ts, -.5, 0))) %>%
  mutate(target_speed = ifelse(target_class == "other", 0, target_speed)) %>% # set target speed = 0 before launch
  mutate(target_class = ifelse(target_class == "other", "visible", target_class)) %>% # join pre-launch with visible  
  mutate(pos_displ = gX - target_x,
         speed_diff = gaze_speed - target_speed) %>%
  mutate(cond_ts = ifelse(target_class == "visible", trial_ts, occ_locked_ts))

vis <- ddata %>%
  filter(target_class == "visible") %>%
  group_by(participant, block, tn_linear, trial_number, trial_dir, success_trial, target_class) %>%
  group_modify(~ resample(.x, start_time = -.5)) %>%
  ungroup 

occ <- ddata %>%
  filter(target_class == "occluded") %>%
  group_by(participant, block, tn_linear, trial_number, trial_dir, success_trial, target_class) %>% 
  filter(n() >= 2) %>%
  group_modify(~ resample(.x, start_time = 0)) %>%
  ungroup 

plotdata <- bind_rows(vis, occ) %>%
  arrange(participant, tn_linear, desc(target_class), newts) %>%
  mutate(newts = round(newts, 4))

```


```{r}
sr <- lindf %>% 
  filter(block != 1) %>%
  distinct(participant, block, tn_linear, success_trial) %>% 
  group_by(participant) %>% 
  summarise(sr = mean(as.numeric(success_trial)-1)) %>%
  ungroup


m <- plotdata %>%
  group_by(participant, target_class, newts) %>%
  filter(n() >= 5) %>%
  group_by(participant, target_class, newts) %>%
  summarise(m_pos_displ = mean(pos_displ, na.rm=T),
         m_speed_diff = mean(speed_diff, na.rm=T),
         n = n()) %>%
  group_by(participant, target_class) %>%
  arrange(newts) %>%
  nest() %>% 
  mutate(data = map(data, ~ .x %>% 
                      mutate(rm_pos_displ = rollapply(m_pos_displ, 12, mean, partial = TRUE, align = "center"),
                             rm_speed_diff = rollapply(m_speed_diff, 12, mean, partial = TRUE, align = "center"))
                    )
         ) %>%
  unnest() %>%
  left_join(sr)

```


```{r}

(tsp <- ddata %>% 
   filter(cond_ts <= 1.5) %>%
  ggplot(aes(cond_ts, pos_displ, group = interaction(participant, tn_linear))) +
  geom_hline(yintercept = 0, alpha = .3, linewidth = 1, linetype = "longdash") + 
  geom_point(alpha = .01) +
  geom_line(aes(newts, rm_pos_displ, color = participant), data = m %>% filter(newts <= 1.5), inherit.aes = F) +
  facet_wrap(~target_class %>% fct_rev, scales = "free_x",
             strip.position = "bottom", 
                labeller = as_labeller(c(visible = "", occluded = "") )) + 
  coord_cartesian(ylim=c(-10, 10)) +
  theme_classic() +
  theme(strip.placement = "outside",
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        strip.background = element_blank()) +
  scale_color_manual(values = sbj_cols) +
  labs(x = "", y = "Positional displacement (deg)", color = "Participant"))


(tss <- ddata %>% 
   filter(cond_ts <= 1.5) %>%
  ggplot(aes(cond_ts, speed_diff, group = interaction(participant, tn_linear))) +
  geom_hline(yintercept = 0, alpha = .3, linewidth = 1, linetype = "longdash") + 
  geom_point(alpha = .01) +
  geom_line(aes(newts, rm_speed_diff, color = participant), data = m %>% filter(newts <= 1.5), inherit.aes = F) +
  facet_wrap(~target_class %>% fct_rev, scales = "free_x",
             strip.position = "bottom", 
                labeller = as_labeller(c(visible = "Time from launch (s)", occluded = "Time from occlusion onset (s)") )) + 
#  ylim(-38,110) +
  coord_cartesian(ylim=c(-38, 110)) +
  theme_classic() +
  theme(strip.placement = "outside",
        strip.background = element_blank()) +
  scale_color_manual(values = sbj_cols) +
  labs(x = "", y = "Speed difference (deg/s)", color = "Participant"))

# scaling should match because all facets have 1.5 s of data
(diff_ts <- ggarrange(tsp, tss, ncol = 1, #widths = c(1, 1), 
               align = "hv", common.legend = T, legend = "right")) 



```


# Fig 3


```{r}
lindf %>%
  filter(target_class %in% c("visible", "occluded"), !(block %in% 2:4 & target_class == "visible")) %>%
  group_by(participant, tn_linear, success_trial, target_class) %>%
  summarise(avg_displ = mean((gX - target_x)),
            sd_displ = sd((gX - target_x)),
            avg_slippage = mean((gaze_speed - target_speed)),
            sd_slippage = sd((gaze_speed - target_speed))) %>%
  group_by(participant, target_class) %>%
  summarise(avg_displ = mean(avg_displ, na.rm=T),
            sd_displ = mean(sd_displ, na.rm=T),
            avg_slippage = mean(avg_slippage, na.rm=T),
            sd_slippage = mean(sd_slippage, na.rm=T),
            sr = ((mean(as.numeric(success_trial))-1)*100) %>% round(1),
            n = n()) %>%
  mutate(across(is.numeric, ~round(.,2)))  %>%
  arrange(target_class, participant)


lindf %>%
  filter(target_class %in% c("visible", "occluded"), !(block %in% 2:4 & target_class == "visible")) %>%
  group_by(participant, target_class) %>%
  summarise(avg_displ = mean((gX - target_x)),
            sd_displ = sd((gX - target_x)),
            avg_slippage = mean((gaze_speed - target_speed)),
            sd_slippage = sd((gaze_speed - target_speed)),
            n = n(),
            n_trials = n_distinct(tn_linear)) %>%
  mutate(across(is.numeric, ~round(.,2)))  %>%
  arrange(target_class, participant)

# means calculated from the "mean line" (not exactly the ones in Fig2, because those have the rolling smoothing)
m %>%
  filter(newts >= 0) %>%
  group_by(target_class, participant) %>%
  summarise(avg_displ = mean(m_pos_displ),
            sd_displ = sd(m_pos_displ),
            avg_slippage = mean(m_speed_diff),
            sd_slippage = sd(m_speed_diff),
            n = n()) %>%
  mutate(across(is.numeric, ~round(.,2)))

```



# Model: Power of 2 (Fexp) + global optimization

(add model comparisons)

```{r}

predicted_accuracy <- function(p_lapse, beta_d, beta_s, fn, df, ...) { 
  p_chance = 1/4
  p_F = fn(df$abs_displ, df$abs_speed_diff, beta_d, beta_s, ...)
  p_correct = (1 - p_chance)*(1-p_lapse)*p_F + p_chance
  p_correct
}

loss_Fexp <- function(p, sbj) {
  
  df = tgt %>% filter(participant == sbj) 
  pred = predicted_accuracy(p_lapse = p[1], beta_d = p[2], beta_s = p[3], fn = Fexp, df = df)  
  error = pred - (as.numeric(df$success_trial) - 1)
  likelihood <- ifelse(df$success_trial == 1,
                       pred,
                       1-pred)
  -sum(log(likelihood))
  
}

Fexp = function(x,y, beta_d, beta_s) exp(-(beta_d*x)**2)*exp(-(beta_s*y)**2) 

```


## DIRECT


```{r}
# initial values
x0 = c(.05, .1, .1)
# lower and upper
lb = c(1.0e-10, 1.0e-10, 1.0e-10)
ub = c(1, 1, 1)
maxeval = 1000 # iterations

get_direct <- function(sbj) {
  
  opts_DIRECT_L <- list("algorithm"= "NLOPT_GN_DIRECT_L",
               "xtol_abs"=1.0e-10,
               "maxeval"=maxeval)
  res_DIRECT_L <- nloptr( x0=x0, 
                  eval_f=loss_Fexp, 
                  lb = lb, 
                  ub = ub, 
                  opts = opts_DIRECT_L,
                 sbj = sbj
                 )
}

dir_res <- lapply(c(1,2,4:10), get_direct)

res_direct <- lapply(dir_res, function(d) data.frame(p_lapse = d$solution[1], beta_d = d$solution[2], beta_s = d$solution[3], loglik = -d$objective)) %>%
  bind_rows() %>%
  mutate(participant = c(1,2,4:10)) %>%
  select(participant, beta_d, beta_s, p_lapse, loglik)

# table of params
res_direct %>%
  mutate(across(3:4, ~round(.,3))) %>%
  mutate(across(c(2,5), ~round(.,2))) 

res_direct %>%
  summarise(across(beta_d:loglik, mean))

```


# Visualization

```{r}
(pred_obs_sbj <- tgt %>%
  left_join(res_direct %>% mutate(participant = factor(participant))) %>%
  mutate(pred = .25 +  (1 - .25 - p_lapse) * Fexp(abs_displ, abs_speed_diff, beta_d, beta_s),
         tq = Fexp(abs_displ, abs_speed_diff, beta_d, beta_s)) %>%
  mutate(dbin = cut(abs_displ, breaks = c(0, 1, 5, 15, 25), include.lowest = T), # cut_number(abs_displ, 4, boundary = 0),
         sbin = cut(abs_speed_diff, breaks = c(0, 2, 30, 300), include.lowest = T) # cut_number(abs_speed_diff, 3, boundary = 0)
           ) %>%
  group_by(dbin) %>% 
  mutate(dmid = (min(abs_displ)+max(abs_displ))/2,
         dmean = mean(abs_displ)) %>%
  group_by(sbin) %>%
  mutate(smid = (min(abs_speed_diff)+max(abs_speed_diff))/2,
         smean = mean(abs_speed_diff)) %>%
  ungroup %>%
  group_by(participant, dbin, sbin, dmid, smid, dmean, smean) %>%
  summarise(mean_pred = mean(pred),
            mean_tq = mean(tq),
            sr = mean(as.numeric(success_trial)-1),
            n = n()) %>%
  gather(var, value, mean_pred, sr) %>% 
  filter(n >= 5, dbin != "(15,25]"
         ) %>% 
  mutate(var = ifelse(var == "mean_pred", "Predicted", "Observed")) %>%
  ggplot(aes(dbin, value, shape = var, color = sbin, group = var)) + 
  geom_point(position = position_dodge(width = .5)) +
  scale_shape_manual(values = 1:2) +
  facet_wrap(~participant) +
  theme_classic() +
  labs(x = "Binned displacement (deg)", y = "Success rate", shape = "", color = "Binned slippage (deg/s)"))

```

```{r}
res_means <- res_direct %>%
  summarise(across(c(p_lapse, beta_d, beta_s), mean))

sr <- tgt %>%
  filter(abs_speed_diff <= 80) %>%
  mutate(dbin = cut(abs_displ, breaks = c(0, .5, 3, 5, 15), include.lowest = T), #cut_number(abs_displ, 5, boundary = 0),
         slippage = cut(abs_speed_diff, breaks = c(0, 2, 10, 30, 80), include.lowest = T) # cut_number(abs_speed_diff, 3, boundary = 0)
           ) %>%
  group_by(dbin, slippage) %>%
  filter(n() >= 10) %>%
  group_by(dbin) %>% # get mid values (& compare to means)
  mutate(dmid = (min(abs_displ)+max(abs_displ))/2,
         dmean = mean(abs_displ)) %>%
  group_by(slippage) %>%
  mutate(smid = (min(abs_speed_diff)+max(abs_speed_diff))/2,
         smean = mean(abs_speed_diff),
         smedian = median(abs_speed_diff),
         smin = min(abs_speed_diff),
         smax = max(abs_speed_diff)) %>%
  ungroup %>%
  group_by(dbin, slippage, dmid, smid, dmean, smean, smedian, smin, smax) %>%
  summarise(sr = mean(as.numeric(success_trial)-1),
            n = n()) %>%
  ungroup()

plines <- sr %>% distinct(slippage, smean, smid, smedian, smin, smax) %>%
  merge(data.frame(abs_displ = seq(0, 10, by = .1))) %>%
  cross_join(res_means) %>%
  mutate(pred = .25 +  (1 - .25 - p_lapse) * Fexp(abs_displ, smean, beta_d, beta_s),
         pred_min = .25 +  (1 - .25 - p_lapse) * Fexp(abs_displ, smin, beta_d, beta_s),
         pred_max = .25 +  (1 - .25 - p_lapse) * Fexp(abs_displ, smax, beta_d, beta_s),
         tq = Fexp(abs_displ, smean, beta_d, beta_s)) 


sr2 <- tgt %>%
  filter(abs_speed_diff <= 80) %>%
  mutate(dbin = cut(abs_displ, breaks = c(0, .5, 3, 5, 15), include.lowest = T), 
         slippage = cut(abs_speed_diff, breaks = c(0, 2, 10, 30, 80), include.lowest = T) 
           ) %>%
  group_by(dbin) %>% # get mid values (& compare to means)
  mutate(dmid = (min(abs_displ)+max(abs_displ))/2,
         dmedian = median(abs_displ),
         dmean = mean(abs_displ)) %>%
  group_by(slippage) %>%
  mutate(smid = (min(abs_speed_diff)+max(abs_speed_diff))/2,
         smedian = median(abs_speed_diff),
         smean = mean(abs_speed_diff)) %>%
  ungroup %>%
  group_by(participant, dbin, slippage, dmid, smid, dmean, smean, dmedian, smedian) %>%
  summarise(sr = mean(as.numeric(success_trial)-1),
            n = n()) %>%
  ungroup()


plines2 <- sr2 %>% distinct(participant, slippage, smean, smedian) %>%
  merge(data.frame(abs_displ = seq(0, 10, by = .1))) %>%
  full_join(res_direct %>% mutate(participant = factor(participant))) %>%
  mutate(pred = .25 +  (1 - .25 - p_lapse) * Fexp(abs_displ, smedian, beta_d, beta_s),
         tq = Fexp(abs_displ, smean, beta_d, beta_s))  


(pred_obs <- sr %>%  
  filter(n >= 10) %>%
  ggplot() + geom_line(aes(abs_displ, pred, color = participant, group = interaction(participant, slippage)), data = plines2, alpha = .7) +
  geom_line(aes(abs_displ, pred), data = plines) +
  geom_point(aes(dmean, sr)) +
  theme_classic() +
  facet_wrap(~slippage, labeller = label_both) +
    xlim(0,10) +
    scale_color_manual(values = sbj_cols) +
    labs(x = "Displacement (deg)", y = "Predicted task accuracy", color = "Participant") +
    scale_y_continuous(breaks = c(.25, .5, .75, 1))
  )


```


